Molecular dynamics simulation of synchronization in driven particles

### 2021-01-02 manuscript/manuscript_ajp_2020.tex
updated fig. 2

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans synchronizing clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
The phenomena 
is often studied with simplified computational models
and can be visualized with %demonstrated
with phase plots or Lissajous figures.
A mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it As we will demonstrate in phase plots in Figs. X, Y, Z})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance a
%pushing a child on a swing at the high point on her path, or
pacemaker (an electrical generator)
pulses to regulate a heart beat
or controlling the pattern of flashing
fireflies with a pulsed light \cite{Agrawal2013}.
These behaviors can be modeled with
numerical studies %can be performed
where
a driving force is applied to simple harmonic system,
either
coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe or
a child on a swing
moving back and forth through equilibrium.  % \cite{}.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold below
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena known as depinning \cite{Reichhardt2017},
and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
The spheres are a few microns in size
and move at rates of microns per second,
so a system can be imaged
with a traditional optical camera
due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction in a seemly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature dependent
phenomena affecting quantum mechanically small systems. \cite{}
[elaborate].

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions.
A circular optical trap is designed so the center is
a local energy minimum where 
a colloid in the beam center is subject to a
uniform distribution of photon bombardment. %, making 
Colloids at a trap edge
experience a net force
due to the uneven 
distribution of photon collisions on its surface.
Depending on the shape of the trap and location of the particle in the trap,
the radiation pressure either pushes colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

We use a classical model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:landscape0}(a).

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
The confining force %arises from a potential function 
 %creating a landscape of potential minima and maxima
%that modify the force on a particle as a function of position
is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The landscape is modulated in the y-direction
with the periodic function 
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the minima
 with simulation units of energy $E_0 = 1$. 
 %We define the substrate strength to be
 %the maximum force the substrate can exert on a particle 
 %$A_p = 2\pi U_0/L$.
 %
 We plot this function in 
 Fig.~\ref{fig:landscape0}
 we use $N_p = 3$ to illustrate the features of
 Eq.~\ref{eq:ysubstrate}.
 In Fig.~\ref{fig:landscape0}(a) we show 
 the $x-y$ plane with a contour plot
 of the the 2D potential energy landscape,
 where 
 the maxima are colored red and the minima colored blue.
 In Fig.~\ref{fig:landscape0}(b) we plot the function
 $U(y)$ to illustrate how the magnitude
 $\vec{F}^{l}_i$ is calculated from particle position in the $y-$direction.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
We illustrate this force in Fig.~\ref{fig:landscape0}(a)
in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
The term overdamped or critically damped indicates
that an oscillator cannot undergo periodic
undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the damping comes from the viscosity of the fluid,
where a sphere of radius $0.5 \mu m$ has a Reynolds number
of $R \approx 10^{-6}$, indicating that linear drag dominates
the particle motion.
See Section~\ref{sec:problems} Exercise~\ref{ex:reynolds} for more information
about the Reynold's number.
In the overdamped model the
linear friction produced by this 
nonconservative force,
$\vec{F}_{drag} = -\eta \vec{v}_i$ is 
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.

@@

We drive a single particle across a periodic landscape 
along the y-direction described in Eq.~\ref{eq:ysubstrate}
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and landscape period $\lambda = 1.825 a_0$.
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$
in simulation units for force. % $E_0/a_0$.

%%HMMM... HAVE YOU INTRODUCED A PHASE SHIFT?

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$, $F^{DC}=0.01$, and $f=0.01$ cycles per time unit.
The time period of the driving force is
$T = 1/f = 100$,
with the initial driving force at is minimum value 
$F^D(0) = F^{DC}$.

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by the spatial period of the washboard as
$y/\lambda$.
The initial particle position is $y=0$,
where the landscape force is minimum $F^l = 0$.
Subsequent substrate

When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape troughs,
and 
the particle hops between minima in the energy landscape.
In Fig.~\ref{fig:0} we demonstrate this relationship,
where over half of a driving period ($T = 50$) %time units
while $F_{D}(t) > 0 $ 
the particle moves
in the positive y-direction
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^D(t)$ where the parameter $F^{DC}$ is ramped to a constant value.  (b) 
  The position as a function of time of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

@@

The hopping pattern of the driven particle is periodic,
and could be achieved over a range of $F^{DC}$.  %, as we will demonstrate in Fig.~\ref{fig:1}
We explore the ranges of periodic hopping patterns
by increasing $F^{DC}$ as a function of time,
as shown in 
%
Fig.~\ref{fig:1}.
%we show the result increasing the 
%force to the particle as a function of time.  
%increased DC
We apply an external applied force, 
with 
a constant $F^{AC}$ with frequency $\omega = 2\pi f$
then slowly increase $F^{DC}$
at a rate of $\Delta F^{DC} = 0.001$ every $\Delta t = 4000$ integer timesteps
or $40$ time units.
%this seems fast, but it is slow enough to estimate
%the velocity accurately.
%this is not be so for a multiple particle system!
By modifying $F^{DC}$ we 
achieve a variety of oscillation modes.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\bar{v}_{y}$ is increasing in non-uniform steps,
with a quantized height of
$\bar{v}_{y} = n \lambda f$,
where $n$ is an integer,
$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
of the landscape,
and $f = 0.01$ cycles per time unit.

@@

>>>>

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to external stimulus \cite{Pikovsky2003}.
In phase oscillations
%Synchronous behaviors where oscillators move in-phase
%(i.e. in time motions)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
is often studied with simple computational models
and can be visualized 
with phase plots or Lissajous figures.
In phase space 
a mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it As we will demonstrate in phase plots in Figs. X, Y, Z})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance a
%pushing a child on a swing at the high point on her path, or
pacemaker (an electrical generator)
pulses to regulate a heart beat
or controlling the pattern of flashing
fireflies with a pulsed light \cite{Agrawal2013}.
These behaviors can be modeled with
numerical studies %can be performed
where
a driving force is applied to simple harmonic system,
either
coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe or
a child on a swing
moving back and forth. % through equilibrium.  % \cite{}.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold below
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena refered to as pinning \cite{Reichhardt2017},
and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
The spheres are a few microns in size
and move at rates of microns per second,
so a system can be imaged
with a traditional optical camera
due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction in a seemly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature and viscosity dependent
phenomena. \cite{}
[elaborate - move to associated problems?].

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions \cite{Ashkin1997}.
A circular optical trap is designed so the center is
a local energy minimum where 
a colloid in the beam center is subject to a
uniform distribution of photon bombardment. %, making 
Colloids at a trap edge
experience a net force
due to the uneven 
distribution of photon collisions on its surface.
Depending on the shape of the trap and location of the particle in the trap,
the radiation pressure either pushes colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as
Shapiro steps in the force-velocity relationship.

@@

We use a classical model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:landscape0}(a).

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The landscape is modulated in the y-direction
with the periodic function 
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the minima
 with simulation units of energy $E_0 = 1$. 
 %We define the substrate strength to be
 %the maximum force the substrate can exert on a particle 
 %$A_p = 2\pi U_0/L$.
 %
 We plot this function in 
 Fig.~\ref{fig:landscape0}
 we use $N_p = 3$ to illustrate the features of
 Eq.~\ref{eq:ysubstrate}.
 In Fig.~\ref{fig:landscape0}(a) we show 
 the $x-y$ plane with a contour plot
 of the the 2D potential energy landscape,
 where 
 the maxima are colored red and the minima colored blue.
 The confining force on a particle $i$
 %arises from a potential function 
 %creating a landscape of potential minima and maxima
 %that modify the force on a particle as a function of position
 is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 In Fig.~\ref{fig:landscape0}(b) we plot the function
 $U(y)$ to illustrate how the magnitude
 $\vec{F}^{l}_i$ is calculated from particle position $y_i$.  
 %in the $y-$direction.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.
We illustrate this force in Fig.~\ref{fig:0}(a).

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
%The term overdamped or critically damped indicates
%that an oscillator cannot undergo periodic
%undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the damping comes from the viscosity of the fluid,
where a sphere of radius $0.5 \mu m$ has a Reynolds number
of $R \approx 10^{-6}$, indicating that linear drag dominates
the particle motion.
See Section~\ref{sec:problems} Exercise~\ref{ex:reynolds} for more information
about the Reynold's number.
In the overdamped model the
linear friction produced by this 
nonconservative force,
$\vec{F}_{drag} = -\eta \vec{v}_i$ is 
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.

@@

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position % as a function of time.
for 
%We drive
a single particle across a periodic landscape. 
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$ force units 
%in simulation units for force. % $E_0/a_0$.
%The substrate is described in 
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and spatial period $\lambda = 1.825 a_0$.
%
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$.
To avoid quenching the system, 
the dc driving force is initially zero   
and 
slowly increased %$F^{DC}$
at a rate of $\Delta F^{DC} = 0.01$ every $\Delta t = 4000$ integer timesteps 
to the constant value $F^{DC}=0.1$,
as shown in the blue curve in Fig.~\ref{fig:0}(a)
where $F^{DC}$ increases for the first $40$ time units.
The combined ac and dc drive $F^D(t)$ is shown in orange,
where 
the magnitude of the ac applied force is initially zero
since $\sin{(0)}=0$.
The vertical dashed lines in panels (a) and (b) coincide
with the condition $\sin{(\omega t)} = 1$,
where the applied force is maximum once $F^{DC}$ has reached its constant value.

%with the initial driving force at is minimum value 
%$F^D(0) = F^{DC}$.

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by the spatial period of the potential %washboard as
$y/\lambda$.
The initial particle position is $y=0$,
where the landscape force is minimum $F^l = 0$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum 
$\sin{(\pi y / \lambda)} = 1 $. 
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
The crossing of the particle position $y$
with the horizontal and vertical lines
indicates the synchronization
of the driving force with the lattice position of the particle,
where driving force is maximum when the particle
is at the landscape minimum.

When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape troughs,
and 
the particle hops between minima in the energy landscape.
{\it UPDATE:} 
In Fig.~\ref{fig:0} we demonstrate this relationship,
where over half of a driving period ($T = 50$) %time units
while $F_{D}(t) > 0 $ 
the particle moves
in the positive y-direction
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^D(t)$ (orange) where the parameter $F^{DC}$ (blue) is ramped to a constant value.  (b) 
  The position as a function of time of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

@@

The hopping pattern of the driven particle is periodic,
and could be achieved over a range of $F^{DC}$.  %, as we will demonstrate in Fig.~\ref{fig:1}
We explore the ranges of periodic hopping patterns
by increasing $F^{DC}$ as a function of time,
as shown in 
%
Fig.~\ref{fig:1}.
%we show the result increasing the 
%force to the particle as a function of time.  
%increased DC
%We apply an external applied force, 
%with 
%a constant $F^{AC}$ with frequency $\omega = 2\pi f$
%then slowly increase $F^{DC}$
%at a rate of $\Delta F^{DC} = 0.001$ every $\Delta t = 4000$ integer timesteps
%or $40$ time units.
%this seems fast, but it is slow enough to estimate
%the velocity accurately.
%this is not be so for a multiple particle system!
The particle 
achieves a variety of oscillation modes as a function of $F^{DC}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\bar{v}_{y}$ is increasing in non-uniform steps,
with a quantized height of
$\bar{v}_{y} = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

@@

     \bibitem{Ashkin1997} Arthur Ashkin, Optical trapping and manipulation of neutral particles using lasers, Proc. Natl. Acad. Sci. U.S.A. 94, 4853–4860 (1997).
